---
title: "HW3 Group5"
author: "Tianyi Zhou, Mduduzi Langwenya, Shengchen Fu, Yanxi Gao, Lin Wang"
date: "2/22/2019"
output: word_document
---
```{r message=FALSE, warning=FALSE}
#load packages
library(tidyverse)
library(knitr)
library(readxl)
library(readr)
library(writexl)
#load revenue data
revenue<-read_csv("VTREVCODE16.csv")
#load ED emergency department data
ED<- read_csv("VTED16.csv")
#load inpatient data
inpatient<-read_csv("VTINP16_upd.csv")
```


###Question 1
####Select main variables from inpatient data to tell the stories
```{r message=FALSE, warning=FALSE}
# change the column Uniq into UNIQ in revenue data
revenue<-revenue %>% rename(UNIQ=Uniq)
#select Unique number, hospital number, admission type, admission source, age, sex, discharge status, primary payer, charges, diagnosis, procedure and patient days columns in inpatient
inpatient_q1<-inpatient%>% select (UNIQ,hnum2,ATYPE,asour,intage,sex,dstat,PPAY,CHRGS,DX1,DX2,DX3,DX4,DX5,DX6,DX7,DX8,PX1,PX2,PX3,pdays) 
#select same columns in ED
ED_q1<-ED %>% select (UNIQ,hnum2,ATYPE,asour,intage,sex,dstat,PPAY,CHRGS,DX1,DX2,DX3,DX4,DX5,DX6,DX7,DX8,PX1,PX2,PX3,pdays) 
## We are doing this part in R because the file is big and it is faster to search for the patient in csv
## When patient went to ER first, their data can be found in both inpatient and ED data. 
#select unique number,revenue code, revenue charges, number of revenue units, CPT code columns in revenue
revenue_q1<-revenue %>% select(UNIQ,REVCODE,REVCHRGS,REVUNITS,CPT)

```

#####Patient 1
```{r message=FALSE, warning=FALSE, results="hide"}
#patient 507033
inpatient_q1 %>%  filter(UNIQ ==507033)
revenue_q1 %>% filter(UNIQ ==507033) %>% group_by(REVCODE) %>% arrange(desc(REVCHRGS))
```

#####Patient 2

```{r message=FALSE, warning=FALSE, results="hide"}
#patient 40436
inpatient_q1 %>%  filter(UNIQ ==40436) 
revenue_q1 %>% filter(UNIQ ==40436) %>% group_by(REVCODE) %>% arrange(desc(REVCHRGS))
```

#####Patient 3

```{r message=FALSE, warning=FALSE, results="hide"}
#patient 859382
ED_q1 %>% filter(UNIQ ==859382)
revenue_q1 %>% filter(UNIQ ==859382) %>% group_by(REVCODE) %>% arrange(desc(REVCHRGS))
```

#####Patient 4

```{r message=FALSE, warning=FALSE, results="hide"}
#patient 1585831
ED_q1 %>%filter(UNIQ ==1585831) 
revenue_q1 %>% filter(UNIQ ==1585831) %>% group_by(REVCODE) %>% arrange(desc(REVCHRGS))
```

#####Patient 5

```{r message=FALSE, warning=FALSE, results="hide"}
#patient 200760
ED_q1 %>% filter(UNIQ ==200760)
revenue_q1 %>% filter(UNIQ ==200760) %>% group_by(REVCODE) %>% arrange(desc(REVCHRGS))
```

#####Patient 6

```{r message=FALSE, warning=FALSE, results="hide"}
#patient 3692
ED_q1 %>% filter(UNIQ ==3692) 
revenue_q1 %>% filter(UNIQ ==3692) %>% group_by(REVCODE) %>% arrange(desc(REVCHRGS))
```

#####Patient 7

```{r message=FALSE, warning=FALSE, results="hide"}
#patient 690326
inpatient_q1 %>%  filter(UNIQ ==690326)
revenue_q1 %>% filter(UNIQ ==690326) %>% group_by(REVCODE) %>% arrange(desc(REVCHRGS))
```

###Question 2
```{r}
#select 3 major insurance payers
num=c(1,2,6,7)
inpatient_major = inpatient %>% filter(PPAY%in%num)
inpatient_major = inpatient_major %>% mutate(Payers = ifelse(PPAY==1,"MEDICARE",
                                     ifelse(PPAY==2,"MEDICAID","Commercial Payers")))
#create table
charge = inpatient_major %>% 
  group_by(Payers, MDC) %>% 
  summarise(Charges=round(sum(CHRGS)/1000000)) %>% 
  drop_na() #drop the rows where at least one column contains NA
cross_tab<-xtabs(Charges ~ Payers + MDC, data = charge)
cross_tab<-as.data.frame.matrix(cross_tab)

#age and sex for major payers
sex = inpatient_major %>% 
  group_by(Payers, sex) %>% 
  summarise(number=n()) %>% 
  drop_na()
age = inpatient_major %>% 
  group_by(Payers, intage) %>% 
  summarise(number=n()) %>% 
  drop_na()
age_sex = inpatient_major %>% 
  group_by(Payers, intage, sex) %>% 
  summarise(number=n()) %>% 
  drop_na()

#export data for piechart and tables
write_xlsx(x = charge, path = "charge.xlsx", col_names = TRUE)
write_xlsx(x = sex, path = "sex.xlsx", col_names = TRUE)
write_xlsx(x = age, path = "age.xlsx", col_names = TRUE)
write_xlsx(x = age_sex, path = "age_sex.xlsx", col_names = TRUE) 
write_xlsx(x = cross_tab, path = "cross_tab.xlsx", col_names = TRUE)
```

###Question 3
#### Reports all the details form the ED file for every emergency department admission that has identified the pt with at least one drug abuse related ICD-10 code 
```{r}
code = c("T40","T41","T42","T43")
drug_abuse<-ED %>% filter_at(vars(starts_with("DX")), any_vars(str_sub(., 1, 3) %in% code))
```

#####1.How many ED visits exactly have been diagnosed as drug user/abuser? 
```{r message=FALSE, warning=FALSE, results="hide"}
length(unique(drug_abuse$UNIQ))
```

#####2.Is the drug use/abuse has been a male problem and that women have much better protection measures to stay away from drug use/abuse let alone overdoses severe enough that lead to an ED admission?  
```{r message=FALSE, warning=FALSE, results="hide"}
drug_abuse %>% group_by(sex) %>% summarise(n=n()) %>%drop_na()
```

#####3.Find the exact dollar amount of the three insurances in Question 2, what was share of each of the total payments?
```{r message=FALSE, warning=FALSE, results="hide"}
sum(drug_abuse$CHRGS)

num=c(1,2,6,7)
drug_abuse_major = drug_abuse %>% filter(PPAY%in%num)
drug_abuse_major = drug_abuse_major %>% mutate(Payers = ifelse(PPAY==1,"MEDICARE",
                                     ifelse(PPAY==2,"MEDICAID","Commercial Payers")))
drug_abuse_major %>% group_by(Payers) %>% summarise(sum=sum(CHRGS),ratio=sum/30741220)
```

#####4.New drugs problem
```{r message=FALSE, warning=FALSE, results="hide"}
syn_narco<-drug_abuse %>% filter_at(vars(starts_with("DX")), any_vars(str_detect(., "^T404")|str_detect(.,"^T4362")))
```

#####5.Name the 3 zip code regions with the highest rate of drug use/abuse. 
```{r message=FALSE, warning=FALSE, results="hide"}
drug_abuse %>% group_by(TXTZIP) %>% summarise(n=n()) %>% arrange(desc(n)) %>% top_n(3)
```

#####6.What are the 10 most common diagnoses of drug use/abuse?
```{r message=FALSE, warning=FALSE, results="hide"}
code = c("T40","T41","T42","T43")
data<-data.frame(drug_abuse[,10:29])
list<-data %>% gather(key,value) %>% group_by(value) %>% tally  %>% filter(str_sub(value, 1, 3) %in% code) %>% arrange(desc(n)) 
list %>% top_n(10)
```


